---
title: "Inference for categorical data"
output:
  pdf_document: default
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
---

In August of 2012, news outlets ranging from the [Washington
Post](http://www.washingtonpost.com/national/on-faith/poll-shows-atheism-on-the-rise-in-the-us/2012/08/13/90020fd6-e57d-11e1-9739-eef99c5fb285_story.html) to the [Huffington
Post](http://www.huffingtonpost.com/2012/08/14/atheism-rise-religiosity-decline-in-america_n_1777031.html)
ran a story about the rise of atheism in America. The source for the story was 
a poll that asked people, "Irrespective of whether you attend a place of 
worship or not, would you say you are a religious person, not a religious 
person or a convinced atheist?" This type of question, which asks people to 
classify themselves in one way or another, is common in polling and generates 
categorical data. In this lab we take a look at the atheism survey and explore 
what's at play when making inference about population proportions using 
categorical data.


```{r libraryload,eval=TRUE}
library(dplyr)

```


## The survey

To access the press release for the poll, conducted by WIN-Gallup 
International, click on the following link:

*<http://www.wingia.com/web/files/richeditor/filemanager/Global_INDEX_of_Religiosity_and_Atheism_PR__6.pdf>*

Take a moment to review the report then address the following questions.

1.  In the first paragraph, several key findings are reported. Do these percentages appear to be *sample statistics* (derived from the data sample) or *population parameters*?

  The several key findings that are reported appears to be from sample statistics which are derived from data sample.


2.  The title of the report is "Global Index of Religiosity and Atheism". To generalize the report's findings to the global human population, what must we assume about the sampling method? Does that seem like a reasonable  assumption?

  Below are the assumptions or conditions
  
  A) The sample should be an simple random sample in each country  and it should be independent of each observation.
  B) It should be less than 10% of global population.
  C) There should be at-least 10 successess and 10 failuers.
  
It satisfies above conditions. So it seems like a reasonable assumption for normal distribution.


## The data

Turn your attention to Table 6 (pages 15 and 16), which reports the
sample size and response percentages for all 57 countries. While this is a useful format to summarize the data, we will base our analysis on the original data set of individual responses to the survey. Load this data set into R with the following command.

```{r head-data, eval=TRUE}
load("more/atheism.RData")
```

3.  What does each row of Table 6 correspond to? What does each row of `atheism` correspond to?

  A) Each row of table 6 corresponds the sample size in individual country and the propotion of religious person, not a religous persion, A convicted athesist and don't know in the taken sample size.
  B) Each row of `atheism`, corresponds to the proption of sample size who is not a religious person.


To investigate the link between these two ways of organizing this data, take a look at the estimated proportion of atheists in the United States. Towards the bottom of Table 6, we see that this is 5%. We should be able to come to the same number using the `atheism` data.


4.  Using the command below, create a new dataframe called `us12` that contains only the rows in `atheism` associated with respondents to the 2012 survey from the United States. Next, calculate the proportion of atheist responses. Does it agree with the percentage in Table 6? If not, why?

```{r us-atheism, eval=TRUE}

us12 <- subset(atheism, nationality == "United States" & year == "2012" )

us12_atheist <- subset(atheism, nationality == "United States" & year == "2012" ) %>% filter(response=="atheist")


#Below percentage matches with Table6

uspercentage <- nrow(us12)/nrow(subset(atheism, nationality == "United States" & year == "2012"))


```

## Inference on proportions

As was hinted at in Exercise 1, Table 6 provides *statistics*, that is, calculations made from the sample of 51,927 people. What we'd like, though, is insight into the population *parameters*. You answer the question, "What proportion of people in your sample reported being atheists?" with a statistic; while the question "What proportion of people on earth would report being atheists" is answered with an estimate of the parameter.

The inferential tools for estimating population proportion are analogous to those used for means in the last chapter: the confidence interval and the hypothesis test.

5.  Write out the conditions for inference to construct a 95% confidence interval for the proportion of atheists in the United States in 2012. Are you confident all conditions are met?

Yes. 

  A) The condition is to get at-least 10 expected successes and failures using the null value. 
  B) The sample size of US is less than 10% of its population.

```{r excercise5,eval=TRUE}

ci <- 1.96
p <- 0.05
se <- sqrt((p*(1-p))/1002)

me <- ci*se

paste("confience interval is (",round(p-me,4),round(p+me,4),")")
```



If the conditions for inference are reasonable, we can either calculate the standard error and construct the interval by hand, or allow the `inference` function to do it for us.

```{r us-atheism-ci, eval=TRUE, tidy = FALSE}
inference(us12$response, est = "proportion", type = "ci", method = "theoretical", success = "atheist")
```

Note that since the goal is to construct an interval estimate for a 
proportion, it's necessary to specify what constitutes a "success", which here is a response of `"atheist"`.

Although formal confidence intervals and hypothesis tests don't show up in the report, suggestions of inference appear at the bottom of page 7: "In general, the error margin for surveys of this kind is +- 3-5% at 95% confidence".


6.  Based on the R output, what is the margin of error for the estimate of the proportion of the proportion of atheists in US in 2012?

Margin of error for the above estimated proportion of athesists in US is +- 1.3%


7.  Using the `inference` function, calculate confidence intervals for the proportion of atheists in 2012 in two other countries of your choice, and report the associated margins of error. Be sure to note whether the conditions for inference are met. It may be helpful to create new data sets for each of the two countries first, and then use these data sets in the `inference` function to construct the confidence intervals.


```{r exercise7,eval=TRUE}

india12 <- subset(atheism, nationality == "India" & year == "2012" ) 

#Validate for atheist in  India
inference(y =india12$response,est="proportion",type="ci", method="theoretical",success = "atheist")

aus12 <- subset(atheism, nationality == "Australia" & year == "2012" ) 

#Validate for atheist in  India
inference(y =aus12$response,est="proportion",type="ci", method="theoretical",success = "atheist")

```


## How does the proportion affect the margin of error?

Imagine you've set out to survey 1000 people on two questions: are you female? 
and are you left-handed? Since both of these sample proportions were 
calculated from the same sample size, they should have the same margin of 
error, right? Wrong! While the margin of error does change with sample size, 
it is also affected by the proportion.

Think back to the formula for the standard error: $SE = \sqrt{p(1-p)/n}$. This 
is then used in the formula for the margin of error for a 95% confidence 
interval: $ME = 1.96\times SE = 1.96\times\sqrt{p(1-p)/n}$. Since the 
population proportion $p$ is in this $ME$ formula, it should make sense that 
the margin of error is in some way dependent on the population proportion. We 
can visualize this relationship by creating a plot of $ME$ vs. $p$.

The first step is to make a vector `p` that is a sequence from 0 to 1 with 
each number separated by 0.01. We can then create a vector of the margin of 
error (`me`) associated with each of these values of `p` using the familiar 
approximate formula ($ME = 2 \times SE$). Lastly, we plot the two vectors 
against each other to reveal their relationship.

```{r me-plot, eval=TRUE}
n <- 1000
p <- seq(0, 1, 0.01)
me <- 2 * sqrt(p * (1 - p)/n)
plot(me ~ p, ylab = "Margin of Error", xlab = "Population Proportion")
```

8.  Describe the relationship between `p` and `me`.

From the formula for me, it is the product of confidence interval and standard error. Standared error is calculated $sqrt(p * (1 - p)/n) $. Also from the about plot, me and p follows a normal distribution.

## Success-failure condition

The textbook emphasizes that you must always check conditions before making 
inference. For inference on proportions, the sample proportion can be assumed 
to be nearly normal if it is based upon a random sample of independent 
observations and if both $np \geq 10$ and $n(1 - p) \geq 10$. This rule of 
thumb is easy enough to follow, but it makes one wonder: what's so special 
about the number 10?

The short answer is: nothing. You could argue that we would be fine with 9 or 
that we really should be using 11. What is the "best" value for such a rule of 
thumb is, at least to some degree, arbitrary. However, when $np$ and $n(1-p)$ 
reaches 10 the sampling distribution is sufficiently normal to use confidence 
intervals and hypothesis tests that are based on that approximation.

We can investigate the interplay between $n$ and $p$ and the shape of the 
sampling distribution by using simulations. To start off, we simulate the 
process of drawing 5000 samples of size 1040 from a population with a true 
atheist proportion of 0.1. For each of the 5000 samples we compute $\hat{p}$ 
and then plot a histogram to visualize their distribution.

```{r sim-np, eval=TRUE}
p <- 0.1
n <- 1040
p_hats <- rep(0, 5000)

for(i in 1:5000){
  samp <- sample(c("atheist", "non_atheist"), n, replace = TRUE, prob = c(p, 1-p))
  p_hats[i] <- sum(samp == "atheist")/n
}


hist(p_hats, main = "p = 0.1, n = 1040", xlim = c(0, 0.18))
```

These commands build up the sampling distribution of $\hat{p}$ using the 
familiar `for` loop. You can read the sampling procedure for the first line of 
code inside the `for` loop as, "take a sample of size $n$ with replacement 
from the choices of atheist and non-atheist with probabilities $p$ and $1 - p$,
respectively." The second line in the loop says, "calculate the proportion of 
atheists in this sample and record this value." The loop allows us to repeat 
this process 5,000 times to build a good representation of the sampling 
distribution.

9.  Describe the sampling distribution of sample proportions at $n = 1040$ and $p = 0.1$. Be sure to note the center, spread, and shape.\
    *Hint:* Remember that R has functions such as `mean` to calculate summary statistics.

```{r excercise9,eval=TRUE}
#Below are summary satistics of sampling distribution of sample proportions
p=0.1
#Centered around 10%
mean(p_hats)

#Standard error - spread

se = sqrt((p*(1-p))/1040)

#95% confidence interval

paste("confience interval is (",round(p-1.96*se,4),",",round(p+1.96*se,4),")")


```



10. Repeat the above simulation three more times but with modified sample
    sizes and proportions: for $n = 400$ and $p = 0.1$, $n = 1040$ and 
    $p = 0.02$, and $n = 400$ and $p = 0.02$. Plot all four histograms
    together by running the `par(mfrow = c(2, 2))` command before creating the 
    histograms. You may need to expand the plot window to accommodate the 
    larger two-by-two plot. Describe the three new sampling distributions. 
    Based on these limited plots, how does $n$ appear to affect the 
    distribution of $\hat{p}$? How does $p$ affect the sampling distribution?


```{r excercise10,eval=TRUE}

#$n = 400$ and $p = 0.1$
p <- 0.1
n <- 400
p4 <- 0.1
n4 <- 1040
p2 <- 0.02
n2 <- 1040
p3 <- 0.02
n3 <- 400

p_hats <- rep(0, 5000)
p_hats2 <- rep(0, 5000)
p_hats3 <- rep(0, 5000)
p_hats4 <- rep(0, 5000)

for(i in 1:5000){
  samp <- sample(c("atheist", "non_atheist"), n, replace = TRUE, prob = c(p, 1-p))
  p_hats[i] <- sum(samp == "atheist")/n
  
  samp <- sample(c("atheist", "non_atheist"), n2, replace = TRUE, prob = c(p2, 1-p2))
  p_hats2[i] <- sum(samp == "atheist")/n2
  
    samp <- sample(c("atheist", "non_atheist"), n3, replace = TRUE, prob = c(p3, 1-p3))
  p_hats3[i] <- sum(samp == "atheist")/n3
  
      samp <- sample(c("atheist", "non_atheist"), n4, replace = TRUE, prob = c(p4, 1-p4))
  p_hats4[i] <- sum(samp == "atheist")/n4
}

par(mfrow = c(2, 2))

hist(p_hats, main = "p = 0.1, n = 400", xlim = c(0, 0.18))
hist(p_hats4, main = "p = 0.1, n = 1040", xlim = c(0, 0.18))
hist(p_hats3, main = "p = 0.02, n = 400", xlim = c(0, 0.18))
hist(p_hats2, main = "p = 0.02, n = 1040", xlim = c(0, 0.18))

par(mfrow = c(1, 1))

```




Once you're done, you can reset the layout of the plotting window by using the 
command `par(mfrow = c(1, 1))` command or clicking on "Clear All" above the 
plotting window (if using RStudio). Note that the latter will get rid of all 
your previous plots.

11. If you refer to Table 6, you'll find that Australia has a sample 
    proportion of 0.1 on a sample size of 1040, and that Ecuador has a sample 
    proportion of 0.02 on 400 subjects. Let's suppose for this exercise that 
    these point estimates are actually the truth. Then given the shape of 
    their respective sampling distributions, do you think it is sensible to 
    proceed with inference and report margin of errors, as the reports does?


```{r excercise11,eval=TRUE}


aus12 <- subset(atheism, nationality == "Australia" & year == "2012" ) 

#Validate for atheist in  Aus
inference(y =aus12$response,est="proportion",type="ci", conflevel = .95,method="theoretical",success = "atheist")

#Ecuador has a sample size of 400.
#success
400*.02
#It is less than 10. So it is not sensible to proceed with inference for Ecuador
```




* * *
## On your own

The question of atheism was asked by WIN-Gallup International in a similar 
survey that was conducted in 2005. (We assume here that sample sizes have 
remained the same.) Table 4 on page 13 of the report summarizes survey results 
from 2005 and 2012 for 39 countries.

-   Answer the following two questions using the `inference` function. As 
    always, write out the hypotheses for any tests you conduct and outline the
    status of the conditions for inference.

    **a.** Is there convincing evidence that Spain has seen a change in its 
    atheism index between 2005 and 2012?\
    *Hint:* Create a new data set for respondents from Spain. Form
    confidence intervals for the true proportion of athiests in both
    years, and determine whether they overlap.


H0: There is no change in atheism index between 2005 and 2012
HA: There is a change in atheism index between 2005 and 2012


```{r onyourown1a,eval=TRUE}

spain12 <- subset(atheism, nationality == "Spain"& year=="2012") 
spain05 <- subset(atheism, nationality == "Spain"& year=="2005") 

#individual ci for 2005
inference(y=spain05$response,est = "proportion",success = "atheist",conflevel = 0.95,alternative = "twosided",type = "ci",method = "theoretical")

#individual ci for 2012
inference(y=spain12$response,est = "proportion",success = "atheist",conflevel = 0.95,alternative = "twosided",type = "ci",method = "theoretical")


#Below percentage matches with Table6

spainper <- nrow(subset(spain05, response=="atheist"))/nrow(spain05)-  nrow(subset(spain12, response=="atheist"))/nrow(spain12)

me <- 1.96*sqrt(((0.10*(1-0.10))/1146)+((0.09*(1-0.09))/1145))


paste("confience interval is (",round(spainper-me,4),",",round(spainper+me,4),")")

```
The above given confidence interval contains 0. So the actual confidence interval cannot be deterimened. We failed to reject null hypotheses.


Conclusion: Spain 2005 & 2012 confidence interval overlaps between each other. So it is hard to say that there is a change or decrease in athesium.


    **b.** Is there convincing evidence that the United States has seen a
    change in its atheism index between 2005 and 2012?



H0: There is no change in United States atheism index between 2005 and 2012
HA: There is a change in United States atheism index between 2005 and 2012


```{r onyourown1,eval=TRUE}

us12 <- subset(atheism, nationality == "United States"& year=="2012") 
us05 <- subset(atheism, nationality == "United States"& year=="2005") 

#individual ci for 2005
inference(y=us05$response,est = "proportion",success = "atheist",conflevel = 0.95,alternative = "twosided",type = "ci",method = "theoretical")

#individual ci for 2012
inference(y=us12$response,est = "proportion",success = "atheist",conflevel = 0.95,alternative = "twosided",type = "ci",method = "theoretical")


#Below percentage matches with Table6

usper <- nrow(subset(us05, response=="atheist"))/nrow(us05)-  nrow(subset(us12, response=="atheist"))/nrow(us12)

me <- 1.96*sqrt(((0.10*(1-0.10))/1146)+((0.09*(1-0.09))/1145))


paste("confience interval is (",round(usper-me,4),",",round(usper+me,4),")")

```

Conclusion:  The above given confidence interval does not contain 0. So there is an increase in atheism. 



-   If in fact there has been no change in the atheism index in the countries 
    listed in Table 4, in how many of those countries would you expect to 
    detect a change (at a significance level of 0.05) simply by chance?\
    *Hint:* Look in the textbook index under Type 1 error.

We would expect there to be a 5% chance of a type I error. 5% of countries will show simply by chance.



-   Suppose you're hired by the local government to estimate the proportion of 
    residents that attend a religious service on a weekly basis. According to 
    the guidelines, the estimate must have a margin of error no greater than 
    1% with 95% confidence. You have no idea what to expect for $p$. How many 
    people would you have to sample to ensure that you are within the 
    guidelines?\
    *Hint:* Refer to your plot of the relationship between $p$ and margin of 
    error. Do not use the data set to answer this question.

If MOE is defined as MOE = 1.96 * sqrt(p*(1-p/n)), we can replace the values as given to compute the sample size.

```{r onyourown3,eval=TRUE}

(1.96^2*(0.5*.5))/0.01^2

```

We need atleast 9604 sample size according to the given guideline.
