---
title: "Week 3 - Eigen Values and Vectors"
author: "Shyam BV"
date: "February 19, 2017"
output:
  html_document:
    code_folding: show
    highlight: textmate
    theme: simplex
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '2'
---


```{r setup, include=FALSE}
if(!require("matlib", character.only = TRUE, quietly = TRUE)) {
  install.packages("matlib")
  library("matlib", character.only = TRUE)
}

if(!require("rgl", character.only = TRUE, quietly = TRUE)) {
  install.packages("rgl")
  library("rgl", character.only = TRUE)
}

if(!require("Matrix", character.only = TRUE, quietly = TRUE)) {
  install.packages("Matrix")
  library("Matrix", character.only = TRUE)
}


```

# Problem Set 1

## (1) What is the rank of the matrix A? 


```{r}

A_4x4 <- matrix(c(1,-1,0,5,2,0,1,4,3,1,-2,-2,4,3,1,-3),nrow=4,ncol=4)

#Option 1:

print(paste("Rank of Matrix:",rankMatrix(A_4x4)[1]))
diag(A_4x4)

#option 2:
factor_output <- function(A) {
  
  L <- diag(nrow =dim(A)[1])
  
for(i in (1:dim(A)[1])) {
  for(j in (1:dim(A)[1])) {
    if(j<i) {
      Eliminate <- diag(nrow =dim(A)[1])

      if(A[i,j]>0) {
      Eliminate[i,j] <- -1*A[i,j]
      }
      else {Eliminate[i,j] <- A[i,j]
      }
              A <- Eliminate %*% A
              
L <- L %*% solve(Eliminate)  

      }
    }
}
print("Lower Matrix with 1 in all pivots. So the rank is 4.")
print(L)

} 


factor_output(matrix(c(1,-1,0,5,2,0,1,4,3,1,-2,-2,4,3,1,-3),nrow=4,ncol=4))


```

## (2) Given an mxn matrix where m > n, what can be the maximum rank? The minimum rank, assuming that the matrix is non-zero? 

For the $mxn$ matrix which has m >n, the maximum rank will be 2. The minimum rank will be 1.

```{r}
A_3x2 <- matrix(c(1,-1,0,5,2,0),nrow=3,ncol=2)

A_3x2_1 <- matrix(c(1,-1,1,-1,1,-1),nrow=3,ncol=2)


print(paste("Maximum Rank of Matrix:",rankMatrix(A_3x2)[1]))

print(paste("Minimum Rank of Matrix:",rankMatrix(A_3x2_1)[1]))

```

## (3) What is the rank of matrix B? 

```{r}

B_3x3 = matrix(c(1,3,2,2,6,4,1,3,2),nrow=3,ncol=3)

print(paste("Maximum Rank of Matrix:",rankMatrix(B_3x3)[1]))


```

Because second column vector is the scalar multiple of first column vector. Third column vector is similar to first column vector.


# Problem Set 2

## Compute the eigenvalues and eigenvectors of the matrix A. You'll need to show your work. You'll need to write out the characteristic polynomial and show your solution. 

```{r}
A_3x3 = matrix(c(1,0,0,2,4,0,3,5,6),nrow=3,ncol=3)
A_3x3

I <- diag(dim(A_3x3)[1])

eigen(A_3x3)

```
$det(\lambda.I_n - A) = 0 $

$$
\left(\begin{array}{cc} 
\lambda-1 & -2 & -3\\
0 & \lambda-4 & 0\\
0 & 0 & \lambda-6\\
\end{array}\right)
=0
$$ 


$(\lambda-1)(\lambda-4)(\lambda-6)-2(0)-3(0)-2(0)-(\lambda-1)(0)-3(0) = 0$

Characteristic Equation: $(\lambda-1)(\lambda-4)(\lambda-6)=0$

Eigen values from above Equation: $\lambda = 1,4,6$


Substitue $\lambda=1$ on A

```{r}

x=1

I <- diag(dim(A_3x3)[1])

A <- x*I-A_3x3


#For row 1,1

  for(i in (2:9)) {
  if(A[1,1]!=1 | A[1,1]!=-1){
    if(A[1,1]/i ==1 | A[1,1]/i ==-1) {
      A[1,] = A[1,]/i
    }
    else if (A[1,1]*i ==1 | A[1,1]*i ==-1) {
      A[1,] = A[1,]*i
    }
  } else {break}
}

#For Row 1,2
  
  for(i in (2:9)) {
  if(A[1,2]!=1 | A[1,2]!=-1){
    if(A[1,2]/i ==1 | A[1,2]/i ==-1) {
      A[1,] = A[1,]/i
    } 
    else if (A[1,2]*i ==1 | A[1,2]*i ==-1) {
      A[1,] = A[1,]*i
    }
   
  } else {break}
}

A[2,] = A[2,]+A[2,2]*A[1,]

#A[2,] = A[2,]*2

  for(i in (2:9)) {
  if(A[2,3]!=1 | A[2,3]!=-1){
    if(A[2,3]/i ==1 | A[2,3]/i ==-1) {
      A[2,] = A[2,]/i
    }
    else if (A[2,3]*i ==1 | A[2,3]*i ==-1) {
      A[2,] = A[2,]*i
    }
  } else {break}
}



A[3,] = A[3,]+A[3,3]*A[2,]
A[1,] = A[1,]+A[1,3]*A[2,]

print("Eigen Vector for eigen value 1:", A)
print(A)


```
$$\lambda_1= 
\left(\begin{array}{cc} 
x_1\\
0\\
0\\
\end{array}\right)
=0
$$ 



```{r}

x=4

I <- diag(dim(A_3x3)[1])

A <- x*I-A_3x3

#For row 1,1

  for(i in (2:9)) {
  if(A[1,1]!=1 | A[1,1]!=-1){
    if(A[1,1]/i ==1 | A[1,1]/i ==-1) {
      A[1,] = A[1,]/i
    }
    else if (A[1,1]*i ==1 | A[1,1]*i ==-1) {
      A[1,] = A[1,]*i
    }
  } else {break}
}


#For Row 1,2
  
  for(i in (2:9)) {
  if(A[1,2]!=1 | A[1,2]!=-1){
    if(A[1,2]/i ==1 | A[1,2]/i ==-1) {
      A[1,] = A[1,]/i
    } 
    else if (A[1,2]*i ==1 | A[1,2]*i ==-1) {
      A[1,] = A[1,]*i
    }
   
  } else {break}
}


A[2,] = A[2,]+A[2,2]*A[1,]

#A[2,] = A[2,]*2

  for(i in (2:9)) {
  if(A[2,3]!=1 | A[2,3]!=-1){
    if(A[2,3]/i ==1 | A[2,3]/i ==-1) {
      A[2,] = A[2,]/i
    }
    else if (A[2,3]*i ==1 | A[2,3]*i ==-1) {
      A[2,] = A[2,]*i
    }
  } else {break}
}


A[3,] = A[3,]+A[3,3]*A[2,]
A[1,] = A[1,]+A[1,3]*A[2,]


print("Eigen Vector for eigen value 4:", A)
print(A)

```

$$\lambda_4= 
\left(\begin{array}{cc} 
0.667x_2\\
x_2\\
0\\
\end{array}\right)
=0
$$ 


```{r}

x=6

I <- diag(dim(A_3x3)[1])

A <- x*I-A_3x3


#For row 1,1

  for(i in (2:9)) {
  if(A[1,1]!=1 | A[1,1]!=-1){
    if(A[1,1]/i ==1 | A[1,1]/i ==-1) {
      A[1,] = A[1,]/i
    }
    else if (A[1,1]*i ==1 | A[1,1]*i ==-1) {
      A[1,] = A[1,]*i
    }
  } else {break}
}


#For row 2,2

  for(i in (2:9)) {
  if(A[2,2]!=1 | A[2,2]!=-1){
    if(A[2,2]/i ==1 | A[2,2]/i ==-1) {
      A[2,] = A[2,]/i
    }
    else if (A[3,3]*i ==1 | A[3,3]*i ==-1) {
      A[2,] = A[2,]*i
    }
  } else {break}
}

A[1,] = A[1,]-A[1,2]*A[2,]


print("Eigen Vector for eigen value 1:", A)
print(A)

```
$$\lambda_6= 
\left(\begin{array}{cc} 
1.6x_3\\
2.5x_3\\
x_3\\
\end{array}\right)
=0
$$ 

