---
title: "Least Squares"
author: "Shyam BV"
date: "March 5, 2017"
output:
  html_document:
    code_folding: show
    highlight: textmate
    theme: simplex
    toc: yes
    toc_float: yes
---


```{r setup, include=FALSE}
if(!require("matlib", character.only = TRUE, quietly = TRUE)) {
  install.packages("matlib")
  library("matlib", character.only = TRUE)
}

if(!require("rgl", character.only = TRUE, quietly = TRUE)) {
  install.packages("rgl")
  library("rgl", character.only = TRUE)
}

if(!require("Matrix", character.only = TRUE, quietly = TRUE)) {
  install.packages("Matrix")
  library("Matrix", character.only = TRUE)

}

if(!require("pracma", character.only = TRUE, quietly = TRUE)) {
  install.packages("pracma")
  library("pracma", character.only = TRUE)
}


```



#1. Problem Set 1

In this problem set we'll work out some properties of the least squares solution that we
reviewed in the weekly readings. Consider the unsolvable system Ax = b as given below:



$$
\left[\begin{array}{cc} 
1 & 0\\
1 & 1\\
1 & 3\\
1 & 4\\
\end{array}\right] 
$$
*

$$
\left[\begin{array}{cc} 
x1\\
x2\\
\end{array}\right] 
$$

=
$$
\left[\begin{array}{cc} 
0\\
8\\
8\\
20\\
\end{array}\right] 
$$

### Write R markdown script to compute $A^T$A and $A^T$b.


```{r}

A = matrix(c(1,1,1,1,0,1,3,4),nrow=4,ncol=2)

b = matrix(c(0,8,8,20),nrow=4,ncol=1)

#A^TA

t(A) %*% A


#A^Tb

t(A) %*% b

```


### Solve for ^x in R using the above two computed matrices.

$\\cap(x)$

```{r}

x_cap = Inverse(t(A) %*% A) %*% (t(A) %*% b)

#x*(x cap) after calculation
x_cap
```


## What is the squared error of this solution?

```{r}

#Projection vector
p = A %*% x_cap

#Error vector. Difference between b and Ax*
e = b - p

#sum of squared error
sum(e*e)


#Function to calculate least squares
lsfit(A, b,intercept=FALSE)$residuals

```


### Instead of b = [0; 8; 8; 20], start with p = [1; 5; 13; 17] and find the exact solution
(i.e. show that this system is solvable as all equations are consistent with each other. This should result in an error vector e = 0).



```{r}
b  = matrix(c(0,8,8,20),nrow=4,ncol=1)
p = matrix(c(1,5,13,17),nrow=4,ncol=1)

x_cap = Inverse(t(A) %*% A) %*% (t(A) %*% b)

#Projection vector
p2 = A %*% x_cap

#Error vector. Difference between p and p2
e = p2 - p

#sum of squared error - This is approximately 0
sum(e*e)

```

### Show that the error e = b - p = [-1;3;-5;3]

```{r}

e = b-p

#error vector between b and p(projection vector)
e

```

### Show that the error e is orthogonal to p and to each of the columns of A.


```{r}
e = b-p


# e is orthogonal to p. If orthogonal p^t * e = 0

t(p) %*% e


# e is orthogonal to A. If orthogonal A^t * e = 0
t(A) %*% e


```

Your code should be able to print all of the above requested quantities. Please include enough comments to make it easy to follow your R markdown document.




# Problem set 2

### Auto Mpg least squares

Consider the modified auto-mpg data (obtained from the UC Irvine Machine Learning
dataset). This dataset contains 5 columns: displacement, horsepower, weight, acceleration,
mpg. We are going to model mpg as a function of the other four variables.

Write an R markdown script that takes in the auto-mpg data, extracts an A matrix
from the first 4 columns and b vector from the fifth (mpg) column. Using the least squares
approach, your code should compute the best fitting solution. That is, and the best fitting
equation that expresses mpg in terms of the other 4 variables. Finally, calculate the fitting
error between the predicted mpg of your model and the actual mpg. Your script should
be able to load in the 5 column data set, extract A and b, and perform the rest of the
calculations. Please have adequate comments in your code to make it easy to follow your
work.

Please complete both problem set 1 & problem set 2 in one R markdown document
and upload it to the site. You don't have to attach the auto-mpg data. Just write your
markdown document in such a way that it expects and loads the auto-mpg data file from the current working directory. As always, your code is expected to compile and run successfully.
Adding test cases to demonstrate that your code is working will be very helpful.



```{r}

auto = read.table("G:/Google_drive/CUNY/Courses/CUNY-repository/605 - Computational Mathematics/Wek 5 - Least Squares/data/auto-mpg.data",header = FALSE)

auto = setNames(auto, c('displacement', 'horsepower','weight', 'acceleration','mpg'))

# Matrix A

A1 = matrix(c(auto$displacement,auto$horsepower,auto$weight,auto$acceleration),nrow=392,ncol=4)

A1=cbind(1,A1)

# Matrix b
b1 = matrix(c(auto$mpg),nrow=392,ncol=1)

#Transpose of A * A

ATA= t(A1) %*% A1 


#Transpose of A * b

ATB= t(A1) %*% b1

#x_cap of the matrix
x_cap = inv(ATA) %*% ATB

#Final fitted line
p = A1 %*% x_cap

#error b- p
e = b1 - p

sum(e)

#validation of residuals(e)
v = lsfit(A1, b1,intercept=FALSE)$residuals

#It is almost equal to zero
sum(e-v)


#Validation via Lm function from statistics
summary(lm(mpg ~ displacement + horsepower + weight + acceleration, data = auto))
x_cap

```
Equation of mpg:

$$mpg = -0.006000871 * displacement - 0.043607731 * horsepower -0.005280508 *weight -0.023147999* acceleration $$


