---
title: "Project 3 - Data Science jobs"
author: "Shyam BV"
date: "October 17, 2016"
output: html_document
---


#Loading necessary packages for analysis

```{r library, eval=TRUE}

knitr::opts_chunk$set(message = FALSE, echo=TRUE)

library(RCurl)
library(dplyr)
library(lubridate)
library(jsonlite)
library(ggplot2)
library(plotly)
library(stringr)
library(tidyr)
library(RMySQL)
library(knitr)
library(DT)
```


```{r dbconnect, eval=TRUE,warning=FALSE,tidy=TRUE, tidy.opts=list(width.cutoff=80)}


#conn <- mysql_dbconn("crime","root","WhiteLotus21")
#conn <- mysql_dbconn('mygroup','mygroup#01','mygroup.c5rotlbjbl71.us-east-1.rds.amazonaws.com','crimedb')



conn <- dbConnect(MySQL(),
                 user = 'mygroup',
                 password = 'mygroup#01',
                 host = 'mygroup.c5rotlbjbl71.us-east-1.rds.amazonaws.com',
                 dbname = 'crimedb')


# List the database tables in crime schema
entdbtables <- dbListTables(conn)

```


## We are fetching the data from JSON file. 

This JSON has two parts. First section has the meta data and second section will have the actual data


```{r sfo-loaddata,eval=TRUE,warning=FALSE,tidy=TRUE, tidy.opts=list(width.cutoff=80)}
#JSON Load

sf_crim.json <- fromJSON("https://raw.githubusercontent.com/bvshyam/Project3_cuny_607/master/Data/sfo_dec2015.json")

#Look for data in JSON
sf_crimedata <-sf_crim.json[['data']]


#Select the necessary columns in the JSON

sf_crime.df <- data.frame(offenseid = sapply(sf_crimedata,function(x) x[[1]]), Category = sapply(sf_crimedata,function(x) x[[10]]), DayOfWeek = sapply(sf_crimedata,function(x) x[[12]]), Date = as.Date(substr(sapply(sf_crimedata,function(x) x[[13]]),1,10),"%Y-%m-%d"),Time =  sapply(sf_crimedata,function(x) x[[14]]),PdDistrict = sapply(sf_crimedata,function(x) x[[15]]),Location= paste(sapply(sf_crimedata,function(x) x[[19]]),",", sapply(sf_crimedata,function(x) x[[18]])), month = month.abb[month(as.Date(substr(sapply(sf_crimedata,function(x) x[[13]]),1,10),"%Y-%m-%d"))], offensearea = sapply(sf_crimedata,function(x) x[[15]]) )


sf_crime_changes<- sf_crime.df 

# Adding Time period for each day

sf_crime_changes <- mutate(sf_crime_changes,Hour= round(as.numeric(str_replace_all(sf_crime_changes$Time,":",".")))) %>%  
  mutate(Date= as.Date(sf_crime_changes$Date,"%m/%d/%Y")) 

sf_crime_changes <- sf_crime_changes %>% mutate(timeframe= sapply(sf_crime_changes$Hour, function(x) { if(x >6 & x <= 11) {"Morning"} else if(x >11 & x <= 16) {"Noon"} else if(x >16 & x <= 20) {"Evening"}  else if(x <=6 | x>20) {"Night" }}))


# Selecing the columns and calculating the additional columns

sf_crime_db <- select(sf_crime_changes,offenseid=offenseid,offense=Category,offensedate=Date,offensetime=Time,dayofweek=DayOfWeek,offensehour=Hour,offensemonth=month,offensearea=offensearea) %>% mutate(citycode="SFO")

```

## Database connection and insert

```{r sfo-datainsert,eval=TRUE,warning=FALSE,tidy=TRUE, tidy.opts=list(width.cutoff=80)}

# Arrange the columns for database insert

sf_crime_db_1 <- select(sf_crime_db,citycode, offenseid, offense, offenselocaldate = offensedate, offenselocaltime = offensetime, dayofweek, offensehour, offensemonth, offensearea)


# Bulk insert from dataframe to MYSQL table called crimedata

#dbWriteTable(conn, value = sf_crime_db_1, name = "crimedata", row.names=F, append = TRUE ) 


```


## Analysis on SFO data

Now we have the required data. We will make some analysis and plot the graphs.


```{r sfo-houroffense,eval=TRUE,warning=FALSE,tidy=TRUE, tidy.opts=list(width.cutoff=80)}

# Select the required columns for comparing hour and offense

sfo_crime_hour <- sf_crime_changes %>% select( Hour, Category) %>% group_by(Hour, Category) %>%  summarise(total =n()) %>% dplyr::arrange(total) 


# Hour vs Offense
    
    p <- ggplot(sfo_crime_hour, aes( x= sfo_crime_hour$Hour,y=sfo_crime_hour$total)) + geom_line(aes(color=Category)) + ggtitle("Hour vs Offense")  + xlab("Hour") +  ylab("Offense")
    ggplotly(p)



# Analyze Crime by offense 

category_count <- table(sf_crime_changes$Category) %>% data.frame() %>% setNames(c("Offense","Frequency"))

datatable(category_count)

#Plot a scattered plot by comparing Offense vs Frequency(occurrence)



a <- list(
  autotick = TRUE,
  tick0 = 0,
  dtick = 1,
  size=24,
    showticklabels = FALSE,
showlegend = FALSE,
  tickcolor = toRGB("white"),
zeroline = FALSE,
  showline = FALSE,
  showticklabels = FALSE,
  showgrid = FALSE,
visible = FALSE,
title = "Offense"
)

b <- list(
title = "Crime count"
)

plot_ly(x=~category_count$Offense,y=~category_count$Frequency,type="scatter",color = category_count$Offense ) %>% layout(xaxis =a,yaxis=b,title = "Offense vs Frequency")




```



#Analyze Crime by Date

```{r sfodatefrequency,eval=TRUE,warning=FALSE,tidy=TRUE, tidy.opts=list(width.cutoff=80)}

#Plot a scattered plot by comparing Offense Date vs Frequency(occurrence)

category_date <- table(sf_crime_changes$Date) %>% data.frame() %>%  setNames(c("Offense_date","Frequency")) 

category_date$Offense_date <- str_replace(category_date$Offense_date,"/2015|2015-","")

datatable(category_date)

#Plot a scattered plot by comparing Offense Date vs Frequency(occurrence)


a1 <- list(
  autotick = TRUE,
  tick0 = 0,
  dtick = 1,
  size=24,
showlegend = FALSE,
tickangle =45,
title=FALSE
)

b1 <- list(
title = "Crime count"
)

plot_ly(x=~category_date$Offense_date,y=~category_date$Frequency,type="area") %>% 
  layout(xaxis =a1,yaxis=b1,title = "Date vs Frequency")


```



#Analyze by timeframe and top crimes


```{r sfo-timefrequency,eval=TRUE,warning=FALSE,tidy=TRUE, tidy.opts=list(width.cutoff=80)}


#Crime according to timeframe and Day

crimesbyday <- table(sf_crime_changes$DayOfWeek,sf_crime_changes$timeframe) %>% data.frame() %>% setNames(c("Day","Day_category","Frequency"))


datatable(crimesbyday)



#Top crimes greater than 5%


mostcrimes <- table(sf_crime_changes$DayOfWeek,sf_crime_changes$Category) %>% data.frame() %>% spread(Var1,Freq) %>% mutate(summary=rowSums(.[,2:8],1))


topcrimes <- mostcrimes[(mostcrimes$summary/sum(mostcrimes$summary))> .05,] %>% data.frame()


datatable(topcrimes)


```




```{r tidy=TRUE, eval=TRUE,warning=FALSE,tidy=TRUE, tidy.opts=list(width.cutoff=80)}
#DATE_FORMAT(offenselocaltime,'%H:%i:%s')

if(isIdCurrent(conn))
{
     allcitycrimedata <-  dbGetQuery(conn,"select crimedata.citycode, offenseid, offense, offenselocaldate, offenselocaltime, dayofweek, offensehour, offensemonth, offensearea from crimedata, city where city.citycode = crimedata.citycode")
    
     
 for (i in seq(1,nrow(allcitycrimedata),1)) 
{    

      if(grepl('LARCENY|ROBBERY|THEFT|STOLEN|EMBEZZLEMENT', allcitycrimedata$offense[i], ignore.case=TRUE ))
     {
           allcitycrimedata$category[i] <- 'LARCENY'
     }
      else if(grepl('RAPE|SEX OFFENSE|SEXUAL|BATTERY|PROSTITUTION', allcitycrimedata$offense[i], ignore.case=TRUE ))
     {
         allcitycrimedata$category[i] <- 'SEX OFFENSE'
     }
      
      else if(grepl('BURGLARY', allcitycrimedata$offense[i], ignore.case=TRUE ))
     {
           allcitycrimedata$category[i] <- 'BURGLARY'
      }
       else if(grepl('MURDER|HOMICIDE', allcitycrimedata$offense[i], ignore.case=TRUE ))
     {
          allcitycrimedata$category[i] <- 'MURDER'
      }
      else if(grepl('NARCOTIC|LIQUOR', allcitycrimedata$offense[i], ignore.case=TRUE ))
     {
          allcitycrimedata$category[i] <- 'NARCOTIC'
      }
      else if(grepl('OTHER|SUICIDE|NON CRIMINAL|NON-CRIMINAL|NON - CRIMINAL|SUSPICIOUS|DECEPTIVE|INTERFERENCE|INTIMIDATION|STALKING|TRESSPASS|SECONDARY|RUNAWAY|MISSING|TRESPASS', allcitycrimedata$offense[i], ignore.case=TRUE ))
     {
       allcitycrimedata$category[i] <- 'OTHER'
     }
     else if( grepl('FORGERY|FRAUD|BAD CHECKS|EXTORTION', allcitycrimedata$offense[i], ignore.case=TRUE ))
      {
           allcitycrimedata$category[i] <- 'FRAUD'
      }
     else if( grepl('PUBLIC INDECENCY|OBSCENITY|DRUNKENNESS|CONDUCT|PUBLIC PEACE|WARRANTS|WEAPON|VEHICLE|LICENSE|DRIVING|BRIBERY|GAMBLING', allcitycrimedata$offense[i], ignore.case=TRUE ))
      {
          allcitycrimedata$category[i] <- 'MISCONDUCT'
      }
       else if( grepl('FAMILY|KIDNAPPING', allcitycrimedata$offense[i], ignore.case=TRUE ))
      {
         allcitycrimedata$category[i] <- 'FAMILY'
      } 
      else if( grepl('CHILDREN', allcitycrimedata$offense[i], ignore.case=TRUE ))
      {
          allcitycrimedata$category[i] <- 'CHILD ABUSE'
       }
       else if( grepl('VANDALISM|DAMAGE|ARSON', allcitycrimedata$offense[i], ignore.case=TRUE ))
      {
          allcitycrimedata$category[i] <- 'VANDALISM'
      }
      else if( grepl('ASSAULT', allcitycrimedata$offense[i], ignore.case=TRUE ))
      {
          allcitycrimedata$category[i] <- 'ASSAULT'
      }
     
 }      # end of for loop
    

}

```

```{r sfo-removetables,eval=TRUE,warning=FALSE,tidy=TRUE, tidy.opts=list(width.cutoff=80)}


#remove SFO tables and variables

remove(sf_crime_db,sf_crime_db_1,sf_crime,sf_day_offense,sfo_crime_hour,topcrimes,sfo_crimedata_hourDF,entdbtables,a,b,a1,b1,category_count,crimesbyday,mostcrimes,sf_crime.df,i,p,sf_crim.json,sf_crimedata)

#Disconnect db

dbDisconnect(conn)

```




