---
title: "Book Recommendation"
author: "Shyam BV"
date: "June 7, 2017"
output:
  html_document:
    highlight: tango
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: yes
      number_sections: yes
      smooth_scroll: no
  pdf_document:
    toc: yes
    toc_depth: '3'
---


###Objective
********

This system recommends different data science books to students in CUNY Class according to their ratings.


```{r include=FALSE}
if(!require("dplyr", character.only = TRUE, quietly = TRUE)) {
  install.packages("dplyr")
  library("dplyr", character.only = TRUE)
}

if(!require("tidyr", character.only = TRUE, quietly = TRUE)) {
  install.packages("tidyr")
  library("tidyr", character.only = TRUE)
}

if(!require("ggplot2", character.only = TRUE, quietly = TRUE)) {
  install.packages("ggplot2")
  library("ggplot2", character.only = TRUE)
}

if(!require("Metrics", character.only = TRUE, quietly = TRUE)) {
  install.packages("Metrics")
  library("Metrics", character.only = TRUE)
}

if(!require("knitr", character.only = TRUE, quietly = TRUE)) {
  install.packages("knitr")
  library("knitr", character.only = TRUE)
}

library(DT)

```

###Dataset
********

Dataset which is used here is from the book ratings which by the stutends of CUNY MSDA program. There are around 12 users and 9 books. Some of them have rated randomly on these books. That data set will be used as to make predictions using prediction metods.


###Loading Data
********

```{r tidy=TRUE, tidy.opts=list(width.cutoff=80)}
ratings_na <- read.csv("data/book_ratings.csv",sep=",",header=TRUE,na.strings = c('Did not read',''))  

datatable(ratings_na,options=list(scrollX='400px'))

#Set temp names
ratings_na <- setNames(ratings_na,c('userId',1:9))

#Remove unused columns
ratings_na <- ratings_na %>% select(-userId)

#backup of full data to calculate RMSE
full_data <- ratings_na


```

###Separate training and test dataset
********

After analyzing the dataset, we can see that the dataset is dense. But there are some missing ratings from each user.


As this is a small dataset, we are allocating 30% to test dataset. Others will be part of training dataset.


```{r}
set.seed(10)

#Random selection of rows from test dataset
test_rows <- sort(sample(nrow(ratings_na),nrow(ratings_na)*.30))

test = ratings_na[test_rows,]

#Selecting only training dataset
ratings_na = ratings_na[-test_rows,]

```


###Raw average of training dataset
********

Calculating raw average of training dataset by converting into a matrix.

```{r}

(train_raw_mean <- mean(as.matrix(ratings_na),na.rm = TRUE))


```


###RMSE calculation
********

Once we calculate the raw average, we can utilize it for calculating RMSE for training and test dataset. 

```{r}
#Create function for RMSE calculation

rmse_calc = function(df,mean_val){

  #As it utilizes NA, converting to 0 to get correct RMSE.
  df[is.na(df)] =0
  
  #Using mean or a dataframe
  if(is.data.frame(df)) mean_val[is.na(mean_val)] =0
  
  return(rmse(df,mean_val))
}


rmse_calc(ratings_na,train_raw_mean)

#Calculate RMSE for test dataset
rmse_calc(test,train_raw_mean)


```



###Bias and Baseline Predictor
********

To calculate baseline predictor, we need to calculate bias. Bias has to be calculated in the prespective of user and books. 

```{r}
#Create function for baseline predictor

baseline_predictor = function(df) {

#User bias: means of each user - raw mean
user_mean =  c(rowMeans(df,na.rm=TRUE)-train_raw_mean)

#book bias: means of each book- raw mean
movie_mean = c(colMeans(df,na.rm=TRUE)-train_raw_mean) 

temp_df = data.frame()

for(i in 1:nrow(df)){
  
  #add all the user and book bias
  final_bias <- train_raw_mean+ user_mean[i] +movie_mean
  temp_df <- rbind(temp_df,final_bias)
  
}

#Set temp names
temp_df = setNames(temp_df, c(1:ncol(temp_df))) %>% data.frame()

#Return the baseline predicted value
return(temp_df)
}

```

### RMSE for baseline predictors
********

Calculate RMSE by comparing baseline predicors and original training, testing and full dataset.


```{r}

#Training dataset RMSE
rmse_calc(ratings_na,baseline_predictor(ratings_na))

#Test dataset RMSE
rmse_calc(test,baseline_predictor(test))

#Full dataset RMSE
rmse_calc(full_data,baseline_predictor(full_data))

```

###Summarize your results
********

Below is the summary of all the baseline predictors

```{r}
#predicted ratings for each user
prediction = baseline_predictor(full_data)

#Setting max and min value

prediction[prediction>5 & !is.na(prediction)] =5
prediction[prediction<1 & !is.na(prediction)] =1

prediction <- setNames(prediction,c("Signal and the Noise","Data Science for Business","Automated Data Collection with R","R for Data Science","Introduction to Statistical Learning","Machine Learning with Scikit-Learn and Tensorflow","Weapons of Math Destruction","Programming Collective Intelligence","R in Action"))


names = c("David Stern","Andy","Walt","Dan Fanelli","James T","Robert Sellers","Tulasi ","Logan","Shyam BV","Yun","Kumudini","Jason Joseph")


datatable(cbind(names,prediction),options=list(scrollX='400px'))
```

After calculating predictions, we will rank the recommended books according to their score.

```{r}
indexes = which(is.na(full_data), arr.ind = TRUE) %>% data.frame() 

indexes['score'] = 0

#Create function for fetching the values

cal_recommendation = function(x){
  
  prediction[x[1],x[2]]
}

indexes['score'] = apply(indexes,1,cal_recommendation)

#Setting Na's to 0
indexes['score'][is.na(indexes['score'])] = 0


names_lookup = data.frame(row = c(1:12), names=c("David Stern","Andy","Walt","Dan Fanelli","James T","Robert Sellers","Tulasi ","Logan","Shyam BV","Yun","Kumudini","Jason Joseph"))

book_lookup = data.frame(col = c(1:9),books = c("Signal and the Noise","Data Science for Business","Automated Data Collection with R","R for Data Science","Introduction to Statistical Learning","Machine Learning with Scikit-Learn and Tensorflow","Weapons of Math Destruction","Programming Collective Intelligence","R in Action"))

full_recommendations = inner_join(indexes,names_lookup) %>% data.frame() %>%  inner_join(book_lookup) %>% arrange(row,-score) %>% group_by(row) %>% mutate(rank=row_number()) %>% select(-col)



datatable(full_recommendations,options=list(scrollY='300px'))
```

