---
title: "Working with API"
author: "Shyam BV"
date: "October 24, 2016"
output:
  html_document:
    fontsize: 17pt
    highlight: pygments
    theme: cerulean
---


####The New York Times web site provides a rich set of APIs, as described here: http://developer.nytimes.com/docs You’ll need to start by signing up for an API key.

####Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it to an R dataframe.

####Below are the libraries which are used. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("tidyjson")

library(tidyjson)
library(dplyr)
library(tidyr)
library(jsonlite)
library(httr)
library(DT)

```


**Once registraion is done with the New York times developer website, we will receive a key to access datasets in developer website. For each section a key needs to be requested.**

```{r moviesdataset,eval=TRUE,warning=FALSE,tidy=TRUE, tidy.opts=list(width.cutoff=80)}

#First let's fetch movie dataset paramenters using GET method

#From API documentation, we can pass order by field


nytimes_movie_get <- GET("https://api.nytimes.com/svc/movies/v2/reviews/all.json",query = list(api_key = "5ca6581c5c0c4d22a89d258eaeb17846",order="by-title",offset=20))

nytimes_movie_get


# Lets convert JSON to dataframe

nytimes_movie_df <- fromJSON("https://api.nytimes.com/svc/movies/v2/reviews/all.json?api_key=5ca6581c5c0c4d22a89d258eaeb17846") %>% data.frame()


head(nytimes_movie_df,1)

```


**In this API call, we will fetch NYtimes symantic dataset**

```{r semanticdataset,eval=TRUE,warning=FALSE,tidy=TRUE, tidy.opts=list(width.cutoff=80)}


#Get Semantic JSON

  nytimes <- GET("https://api.nytimes.com/svc/semantic/v2/concept/search.json",query = list(api_key = "5ca6581c5c0c4d22a89d258eaeb17846",query="Active")) 

  # Lets convert JSON to dataframe 

  
nytimes.semantic.df <- fromJSON("https://api.nytimes.com/svc/semantic/v2/concept/search.json?api_key=5ca6581c5c0c4d22a89d258eaeb17846&query=Active") %>%  data.frame()

head(nytimes.semantic.df,2)



```

###Presidential  candidates dataset from propublica API (https://propublica.github.io/campaign-finance-api-docs/#presidential-candidate-totals)


```{r congressapi,eval=TRUE,warning=FALSE,tidy=TRUE, tidy.opts=list(width.cutoff=50)}

#Keys are passed to header of the URL

key <- c("M28yQdX0TI2adVJZmO2yu8T1dnrw5hZfZ9z15Cqi")
names(key) <- "X-API-Key"

presidential_get <- GET("https://api.propublica.org/campaign-finance/v1/2016/president/totals.json", add_headers(.headers = key),content_type_json()) 

#Convert in into JSON
presidential_get.df <- content(presidential_get) %>% .$results %>% toJSON() %>% fromJSON()

head(presidential_get.df,2)


```

********

###Full contact API from website https://www.fullcontact.com/developer/

This website can be used to fetch many information of persons via his email and phonenumber. 


```{r fullcontactkey,include=FALSE}
fullkey <- c("dd3032a4fca36eb5")

```


**Likelihood of a person presence according fullcontact**

```{r fullcontact,eval=TRUE}

#Get key from fullcontact
names(fullkey) <- "X-FullContact-APIKey"

#Full contact info likelihood

fullcontact_get.df.new <- GET("https://api.fullcontact.com/v2/name/normalizer.json?q=Hillary%20Clinton",add_headers(.headers =fullkey),content_type_json()) %>% content(.) %>% toJSON() %>% fromJSON()

paste("Likelihood of the person is ",fullcontact_get.df.new$likelihood*100,"%")


```



