---
title: "Assignment 2 - Traces and Determinants"
author: "Shyam BV"
date: "February 12, 2017"
output:
  html_document:
    code_folding: show
    highlight: textmate
    theme: simplex
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '2'
---

```{r setup, include=FALSE}
if(!require("matlib", character.only = TRUE, quietly = TRUE)) {
  install.packages("matlib")
  library("matlib", character.only = TRUE)
}

if(!require("rgl", character.only = TRUE, quietly = TRUE)) {
  install.packages("rgl")
  library("rgl", character.only = TRUE)
}

if(!require("Matrix", character.only = TRUE, quietly = TRUE)) {
  install.packages("Matrix")
  library("Matrix", character.only = TRUE)
}

```


#Problem 1

##Part 1
**(1) Show that $A^TA != AA^T$ in general. (Proof and demonstration.)** 

Here lets prove the above condition with examples and function


```{r}


output <- function(A) {
  
LHS <- t(A)  %*% A

RHS <- A %*% t(A)

print(LHS)
print(RHS)
return (LHS==RHS)
  
}

# 2x2 matrix
output(matrix(c(1,3,2,1),2))


# 3x3 matrix
output(matrix(c(1,2,3,1,1,1,2,0,1),nrow=3))

#4x4
output(matrix(c(1,2,1,3,4,1,2,3,1,1,5,7,2,5,6,2),nrow=4))
```

Above output shows that $A^TA != AA^T$. Generally $A.B != B.A$


##Part 2

**(2) For a special type of square matrix A, we get A T A = AA T . Under what conditions could this be true? (Hint: The Identity matrixIis an example of such a matrix). **

But the above mentioned case is not true in all the places. Below are some exceptional scenarios.

1. It happens when the elements above and below the diagonal are equal. 
2. Also the row order can be altered by Permutation matrix. After alteration the matrix should have symmentry
```{r}

# 3x3 Identity matrix
output(matrix(c(1,0,0,0,1,0,0,0,1),nrow=3))

# 2x2 Symmetric matrix
output(matrix(c(1,2,2,7),nrow=2,ncol=2))

# 3x3 Symmetric matrix
output(matrix(c(1,4,5,4,2,6,5,6,3),nrow=3,ncol=3))


```


#Problem 2

**Matrix factorization is a very important problem. There are supercomputers built just to do matrix factorizations. Every second you are on an airplane, matrices are being factorized. Radars that track  ights use a technique called Kalman  filtering. At the heart of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear systems of equations when they track your  fight using radars. **

**Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer. Please submit your response in an R Markdown document using our class naming convention, E.g.  LFulton_Assignment2_PS2.png.**

**You don't have to worry about permuting rows of A and you can assume that A is less than 5x5, if you need to hard-code any variables in your code. If you doing the entire assignment in R, then please submit only one markdown document for both the problems. **



```{r}

factor_output <- function(A) {
  
  L <- diag(nrow =dim(A)[1])
  
for(i in (1:dim(A)[1])) {
  for(j in (1:dim(A)[1])) {
    if(j<i) {
      Eliminate <- diag(nrow =dim(A)[1])
      lower <- diag(nrow =dim(A)[1])

      if(A[i,j]>0) {
      Eliminate[i,j] <- -1*A[i,j]
      }
      else {Eliminate[i,j] <- A[i,j]
      }
              A <- Eliminate %*% A
              
L <- L %*% solve(Eliminate)  

      }
    }
}
print("Lower Matrix")
print(L)

print("Upper Matrix")
print(A)

print("Multiply (L*U)")
print(L %*% A)
#remove(L)

} 


factor_output(matrix(c(1,2,3,1,1,1,2,0,1),nrow=3,ncol=3))



```

