---
title: "Kickstarter Projects"
author: "Shyam BV"
date: "November 7, 2016"
output: html_document
---


```{r include=FALSE, cache=FALSE}
# DO NOT REMOVE
# THIS IS FOR SETTING SOME PLOTTING PARAMETERS SO THAT YOUR PLOTS DON'T TAKE UP TOO MUCH SPACE
# IF YOU WOULD LIKE TO CHANGE THESE, SEE HELP FILES ON THE par() FUNCTION
# OR ASK FOR HELP
library(dplyr)
library(knitr)
## set global chunk options
opts_chunk$set(fig.path='figure/manual-', cache.path='cache/manual-', fig.align='center', fig.show='hold', par=TRUE)
## tune details of base graphics (http://yihui.name/knitr/hooks)
knit_hooks$set(par=function(before, options, envir){
if (before && options$fig.show!='none') par(mar=c(4,4,.2,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
})
```



```{r loaddata,eval=TRUE}
# load data of 2016 Q1

kickstarter <- read.csv ("./data/kickstarter_projects.csv",header = TRUE, sep = ",",stringsAsFactors = FALSE,na.strings = c("Null","NULL","n/a","na")) %>% select ( id,	kickstarter_id,	name,	blurb,main_category,	sub_category,	launched_at,	deadline,	state,	goal,	pledged,	percentage_funded,	backers_count,	became_overachiever_at,		underdog_chosen_by_user_id,	photo,	web_url,	currency)


summary(kickstarter)
```


### Research question 

**You should phrase your research question in a way that matches up with the scope of inference your dataset allows for.**

1. Are technology focussed companies more likely to fail?
2. What is the projected funding for all the projects in 2016?
3. Is success rate is directly related to the goal amount?
4. Will the goal be achieved if the project has more days?
5. Confidence interval of the goal amount


### Cases 

**What are the cases, and how many are there?**



https://drive.google.com/file/d/0B1fUxQl4wYFrM0M4YUxqSVlIVVk/view