---
title: "Lending Loans"
author: "Shyam BV"
date: "November 7, 2016"
output: html_document
---



```{r test,include=FALSE, cache=FALSE}
# DO NOT REMOVE
# THIS IS FOR SETTING SOME PLOTTING PARAMETERS SO THAT YOUR PLOTS DON'T TAKE UP TOO MUCH SPACE
# IF YOU WOULD LIKE TO CHANGE THESE, SEE HELP FILES ON THE par() FUNCTION
# OR ASK FOR HELP
library(dplyr)
library(knitr)
library(stringr)
library(ggplot2)
library(plotly)

load("data/nc.RData")

## set global chunk options
opts_chunk$set(fig.path='figure/manual-', cache.path='cache/manual-', fig.align='center', fig.show='hold', par=TRUE)
## tune details of base graphics (http://yihui.name/knitr/hooks)
knit_hooks$set(par=function(before, options, envir){
if (before && options$fig.show!='none') par(mar=c(4,4,.2,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
})

normalPlot <- function(mean=0, sd=1, bounds=c(-1,1), tails=FALSE) {
	x <- seq(-4,4,length=100)*sd + mean
	hx <- dnorm(x,mean,sd)

	plot(x, hx, type="n", xlab="x-Axis", ylab="",
		 main="Normal Distribution", axes=FALSE)
	lines(x, hx)

	if(tails) {
		i.low <- x <= bounds[1]
		i.high <- x >= bounds[2]
		polygon(c(x[i.low],bounds[1]), c(hx[i.low], 0), col="red")
		polygon(c(bounds[2],x[i.high]), c(0,hx[i.high]), col="red")
	} else {
		i <- x >= bounds[1] & x <= bounds[2]
		polygon(c(bounds[1],x[i],bounds[2]), c(0,hx[i],0), col="red")
		area <- pnorm(bounds[2], mean, sd) - pnorm(bounds[1], mean, sd)
		if(diff(bounds) > 0) {
			result <- paste("P(",bounds[1],"< x <",bounds[2],") =",
							signif(area, digits=3))
			mtext(result,3)
		}
	}
	axis(1)
}


```



```{r loaddata,eval=TRUE}
# load data of 2016 Q1

set.seed(7340)

#Getting the sample size of 1000 from loans stats 2016 Q1 dataset 

loans_sample <- read.csv ("C:/Users/Shyam/Google Drive/CUNY/Assignment_Repositories/606 - Project/LoanStats_securev1_2016Q1.csv",header = TRUE,skip = 1, sep = ",",stringsAsFactors = FALSE,skipNul = TRUE) %>% sample_n(1000,replace = FALSE)

#loans_sample <- read.csv ("C:/Users/paperspace/Google Drive/CUNY/Assignment_Repositories/606 - Project/LoanStats_securev1_2016Q1.csv",header = TRUE,skip = 1, sep = ",",stringsAsFactors = FALSE,skipNul = TRUE) %>% sample_n(1000,replace = FALSE)



# 1. Select the required columns
# 2. Convert the column term to numeric
# 3. Remove unwanted columns and change the NA data(if any)
# 4. Calculate the total payment by each customer adding the interest rate

loans_summary <-  
loans_sample %>% select(c(id,loan_amnt,funded_amnt,term,int_rate,emp_length,home_ownership,annual_inc,verification_status,addr_state,fico_range_low,fico_range_high,issue_d,loan_status,purpose,total_pymnt_inv)) %>% 
  mutate(fico_score=(fico_range_low+fico_range_high)/2) %>% 
    mutate(term1=as.numeric(str_trim(str_replace(term,"months","")))) %>% 
      select(-term,-fico_range_low,-fico_range_high) %>% 
        mutate(emp_length=str_replace(emp_length,c("n/a"),"NA")) %>% 
          mutate(int_rate=(as.numeric(str_replace(as.character(int_rate),"%",""))))  %>%  mutate(total_pymnt=(funded_amnt+(funded_amnt*int_rate*(term1/12))/100))



#Change interest rate to numeric

loans_summary$int_rate <- as.numeric(loans_summary$int_rate)

#Changing required columns as factors and numeric

loans_summary$home_ownership <- as.factor(loans_summary$home_ownership)

loans_summary$verification_status <- as.factor(loans_summary$verification_status)

loans_summary$purpose <- as.factor(loans_summary$purpose)
loans_summary$loan_status <- as.factor(loans_summary$loan_status)

#As we are concentrated on home ownership, we are replacing OWN to 

```


### Research question 

**You should phrase your research question in a way that matches up with the scope of inference your dataset allows for.**

Below are some research questions from Sample data

1. Is loan interest % predictive of credit score?
2. Is loan funded amount are equal for different purpose of loan request?
3. Does different states have equal interest rate?
4. Does home ownership really impact FICO scores or just by chance?
5. Number of loan requests which will be received in next month


### Cases 

**What are the cases, and how many are there?**

For 2016 Q1, there are around 133K observations. For modelling purposes, I have taken a sample of 1000 rows. 


### Data collection 

**Describe the method of data collection.**

Data is collected from Lending loan website (https://www.lendingclub.com/info/download-data.action). 


### Type of study 

**What type of study is this (observational/experiment)?**

This is an observational study.


### Data Source 

**If you collected the data, state self-collected. If not, provide a citation/link.**

Link: https://www.lendingclub.com/info/download-data.action

*******

### Response 

**What is the response variable, and what type is it (numerical/categorical)?**

Response provided

### Explanatory 

**What is the explanatory variable, and what type is it (numerical/categorical)?**


Below table summarizes the question, response and explanatory variable. It also shows whether it is Numerical or Categorical.


Question | Response Variable | Explanatory Variable
-------------| -------------|-----------------
1. Is loan interest rate % predictive of credit score? | Interest rate % (Numerical)  | Credit score (Numerical) 
2. Is loan funded amount are equal for different purpose of loan request? | Funded amount (Numerical) | Different purpose (Categorical) 
3. Does different states have equal interest rate? | Interest rate % (Numerical) | States (Categorical)
4. Does home ownership really impact FICO scores or just by chance? | FICO scores (Numerical) | Home Ownership (Categorical)
5. Number of loan requests which will be received in next month | Loan requests (Numerical) | Months (Categorical)


### Relevant summary statistics 

**Provide relevant to your research question. For example, if you’re comparing means across groups provide means, SDs, sample sizes of each group. This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

Sample size = 1000

```{r summarystatistic,eval=TRUE}

summary(loans_summary)

loandatamean <- group_by(loans_summary,purpose) %>% summarise(mean(funded_amnt))

loandatasd <- group_by(loans_summary,purpose) %>% summarise(sd(funded_amnt))

ggplot(loans_summary,aes(x=funded_amnt)) + geom_histogram()


```



```{r charts,eval=TRUE}



#Histogram of funded amount

ggplot(loans_summary,aes(x=funded_amnt)) + geom_histogram()

#Histogram of Interest rate %

ggplot(loans_summary,aes(x=int_rate)) + geom_histogram()


#fico score vs interest rate
ggplotly(ggplot(loans_summary,aes(x=fico_score,y=int_rate)) + geom_point(aes(color=purpose),size=1,alpha=0.8) + geom_smooth(method = "lm"))


#funded amount are equal for different purpose

ggplot(loans_summary,aes(x=purpose,y=funded_amnt)) + geom_boxplot()


#Interest rate vs funded amount

ggplot(loans_summary,aes(x=int_rate,y=funded_amnt)) + geom_point(aes(color=term1)) 



# State vs  Interest rate % 

state_interestrate <- loans_summary %>% select(addr_state,int_rate) %>%  group_by(addr_state) %>% summarise(avg =mean(int_rate))

a1 <- list(autotick = TRUE, title = "State")

b1 <- list(title = "Interest rate %")

plot_ly(x = ~state_interestrate$addr_state, y = ~state_interestrate$avg, type = "area") %>% layout(xaxis=a1,yaxis=b1,title="State vs Average interest rate")


#Home ownership vs FICO scores

ggplot(loans_summary,aes(x=fico_score)) + geom_histogram() + facet_grid(.~home_ownership)


#Loan request vs month

plot_ly(x = ~loans_summary$issue_d)

#remove(loans_sample,loans_summary,loandatamean,loandatasd,a1,p2,p3,p4,p5,a1,b1,g,h)

```


Confidence interval of the funded amount

```{r confidence,eval=TRUE}

#Point estimate for the population funded amount

p = mean(loans_summary$funded_amnt)
n=1000

se = (sd(loans_summary$funded_amnt))/sqrt(1000)

# 95% confidence interval

me = 1.96*se

# Confidence interval of funded amount

normalPlot(p,se,bounds=c(p-me,p+me))

```


Hypothesis test between the funded amount of RENT vs MORTGAGE

```{r hypothesistest, eval=TRUE}

loan_summary1 <- filter(loans_summary,home_ownership==c("RENT","MORTGAGE"))

loan_summary1$home_ownership <- as.character(loan_summary1$home_ownership)


#Hypothesys test
inference(y=loan_summary1$funded_amnt,x=loan_summary1$home_ownership,est="mean",null=0,alternative="twosided",type="ht",method="theoretical")

# As the p-value is low, we reject the hypothesys
```

```{r confidencetest1, eval=TRUE}


#Confidence interval test
inference(y=loan_summary1$funded_amnt,x=loan_summary1$home_ownership,est="mean",null=0,alternative="twosided",type="ci",conflevel=0.95,method="theoretical")

```

