---
title: "Bias Variance - Tradeoff"
author: "Shyam BV"
date: "May 4, 2017"
output: html_document
---

```{r include=FALSE}
library(stats)
library(boot)

```

#####Load auto data

```{r}
#auto data

auto = read.table("./data/auto-mpg.data",header = FALSE)

auto = setNames(auto, c('displacement', 'horsepower','weight', 'acceleration','mpg'))
head(auto)
```

#####K-fold

In k-fold cross-validation, the original sample is randomly partitioned into k equal sized subsamples. Of the k subsamples, a single subsample is retained as the validation data for testing the model, and the remaining k − 1 subsamples are used as training data. The cross-validation process is then repeated k times (the folds), with each of the k subsamples used exactly once as the validation data. And the average of K samples is the output of MSE.

```{r}
#Set seed
set.seed(7340)
cv.err5 = NULL

#Loop until 8 polynomial

for(i in 1:8) {
glm.fit = glm(mpg ~ poly(displacement + horsepower + weight + acceleration,i), data = auto)

cv.err5[i] = cv.glm(auto, glm.fit, K =5)$delta[1]

}

cv.err5

#Plot the mse for each polynomial
degree=1:i 
plot(degree,cv.err5,type='b')

```


#####Leave-one-out cross-validation (LOOCV) method

```{r}
#Leave-one-out cross-validation method

glm.fit.loocv = glm(mpg ~ displacement + horsepower + weight + acceleration, data = auto)

cv.err = cv.glm(auto, glm.fit.loocv)$delta[1]

#MSE of LOOCV
cv.err
```